---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(0)
```

# distributional

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![R build status](https://github.com/mitchelloharawild/distributional/workflows/R-CMD-check/badge.svg)](https://github.com/mitchelloharawild/distributional)
[![Coverage Status](https://codecov.io/gh/mitchelloharawild/distributional/branch/master/graph/badge.svg)](https://codecov.io/github/mitchelloharawild/distributional?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/distributional)](https://CRAN.R-project.org/package=distributional)
<!-- badges: end -->

The distributional package allows distributions to be used in a vectorised context. It provides methods which are minimal wrappers to the standard d, p, q, and r distribution functions which are applied to each distribution in the vector. Additional distributional statistics can be computed, including the `mean()`, `median()`, `variance()`, and intervals with `hilo()`. 

The distributional nature of a model's predictions is often understated, with defaults of `predict()` methods usually only producing point predictions. The `forecast()` function from the [forecast package](https://CRAN.R-project.org/package=forecast) goes further in illustrating uncertainty by producing point forecasts and intervals by default, however the user's ability to interact with them is limited. This package vectorises distributions and provides methods for working with them, making entire distributions suitable prediction outputs for model functions.

## Installation

You can install the released version of distributional from [CRAN](https://CRAN.R-project.org/package=distributional) with:

```{r, eval = FALSE}
install.packages("distributional")
```

The development version can be installed from [GitHub](https://github.com/mitchelloharawild/distributional) with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("mitchelloharawild/distributional")
```

## Examples

Distributions are created using `dist_*()` functions.
Currently only the normal distribution is supported for testing purposes.

```{r object}
library(distributional)
my_dist <- c(dist_normal(mu = 0, sigma = 1), dist_student_t(df = 10))
my_dist
```

The standard four distribution functions in R are usable via these generics:

```{r}
density(my_dist, 0) # c(dnorm(0, mean = 0, sd = 1), dt(0, df = 10))
cdf(my_dist, 5) # c(pnorm(5, mean = 0, sd = 1), pt(5, df = 10))
quantile(my_dist, 0.1) # c(qnorm(0.1, mean = 0, sd = 1), qt(0.1, df = 10))
generate(my_dist, 10) # list(rnorm(10, mean = 0, sd = 1), rt(10, df = 10))
```

You can also compute intervals using `hilo()`

```{r hilo}
hilo(my_dist, 0.95)
```

Additionally, some distributions may support other methods such as mathematical operations and summary measures. If the methods aren't supported, a transformed distribution will be created.

```{r math}
my_dist
my_dist*3 + 2
mean(my_dist)
variance(my_dist)
```

You can also visualise the distribution(s) using `autoplot()`.

```{r plot}
library(ggplot2)
autoplot(my_dist, type = "pdf")
autoplot(my_dist, type = "cdf")
```

## Related work

There are several packages which unify interfaces for distributions in R:

* stats provides functions to work with possibly multiple distributions (comparisons made below).
* [distributions3](https://cran.r-project.org/package=distributions3) represents singular distributions using S3, with particularly nice documentation.
* [distr](https://cran.r-project.org/package=distr) represents singular distributions using S4.
* [distr6](https://cran.r-project.org/package=distr6) represents singular distributions using R6.
* Many more in the [CRAN task view](https://cran.r-project.org/view=Distributions)

This package differs from the above libraries by storing the distributions in a vectorised format. It does this using [vctrs](https://vctrs.r-lib.org/), so it should play nicely with the tidyverse (try putting distributions into a tibble!).
