---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(0)
```

# distributional

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/distributional)](https://CRAN.R-project.org/package=distributional)
[![R build status](https://github.com/mitchelloharawild/distributional/workflows/main/badge.svg)](https://github.com/mitchelloharawild/distributional)
<!-- badges: end -->

The distributional package allows distributions to be used in a data context.

## Installation

<!-- You can install the released version of distributional from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ```{r, eval = FALSE} -->
<!-- install.packages("distributional") -->
<!-- ``` -->

The development version can be installed from [GitHub](https://github.com/mitchelloharawild/distributional) with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("mitchelloharawild/distributional")
```
## Example

Distributions are created using `dist_*()` functions.
Currently only the normal distribution is supported for testing purposes.

```{r object}
library(distributional)
my_dist <- dist_normal(mu = 1:10, sigma = 1)
my_dist
```

The resulting object is a [vctrs](https://vctrs.r-lib.org/) vector, so it should play nicely with the tidyverse.

```{r tidyverse}
library(tibble)
tibble(my_dist)
```

The standard four distribution functions in R are usable via these generics:

```{r}
density(my_dist, 0) # dnorm(0, mean = 1:10, sd = 1)
cdf(my_dist, 5) # pnorm(5, mean = 1:10, sd = 1)
quantile(my_dist, 0.1) # qnorm(0.1, mean = 1:10, sd = 1)
generate(my_dist, 10) # lapply(1:10, function(mu) rnorm(10, mean = mu, sd = 1))
```

You can also compute intervals using `hilo()`

```{r hilo}
hilo(my_dist, 0.95)
```

Additionally, some distributions may support other methods such as mathematical operations and summary measures.

```{r math}
my_dist
my_dist*3 + 2
mean(my_dist)
```
