---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(0)
```

# distributional

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/mitchelloharawild/distributional/workflows/ghci/badge.svg)](https://github.com/mitchelloharawild/distributional)
[![Coverage Status](https://codecov.io/gh/mitchelloharawild/distributional/branch/master/graph/badge.svg)](https://codecov.io/github/mitchelloharawild/distributional?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/distributional)](https://CRAN.R-project.org/package=distributional)
<!-- badges: end -->

The distributional package allows distributions to be used in a data context.

## Installation

<!-- You can install the released version of distributional from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ```{r, eval = FALSE} -->
<!-- install.packages("distributional") -->
<!-- ``` -->

The development version can be installed from [GitHub](https://github.com/mitchelloharawild/distributional) with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("mitchelloharawild/distributional")
```

## Brief background

The [fabletools](https://github.com/tidyverts/fabletools) package produces distributional forecasts. Returning distributions for predictions isn't particularly new (see `predict.lm()` for instance), however it is unique in how it represents the distributions. Unlike other packages which return the distribution's parameters (expecting users to use the parameters appropriately), fabletools returns a classed distribution object.

This package aims to generalise the crude distribution objects developed for fabletools, into light-weight and general distribution classes for use in prediction outputs of models. There are several packages which implement distributions for R:

* stats provides functions to work with possibly multiple distributions (comparisons made below).
* [distributions3](https://cran.r-project.org/package=distributions3) represents singular distributions using S3, with particularly nice documentation.
* [distr](https://cran.r-project.org/package=distr) represents singular distributions using S4.
* [distr6](https://cran.r-project.org/package=distr6) represents singular distributions using R6.
* Many more in the [CRAN task view](https://cran.r-project.org/view=Distributions)

This package differs from the above libraries by storing the distributions in a vectorised format. It does this using [vctrs](https://vctrs.r-lib.org/), so it should play nicely with the tidyverse (try putting distributions into a tibble!).

Additionally, this package aims to support modifiers on distributions such as transformations (such as `log`), zero-inflation, and truncation.

## Example

Distributions are created using `dist_*()` functions.
Currently only the normal distribution is supported for testing purposes.

```{r object}
library(distributional)
my_dist <- dist_normal(mu = 1:10, sigma = 1)
my_dist
```

The standard four distribution functions in R are usable via these generics:

```{r}
density(my_dist, 0) # dnorm(0, mean = 1:10, sd = 1)
cdf(my_dist, 5) # pnorm(5, mean = 1:10, sd = 1)
quantile(my_dist, 0.1) # qnorm(0.1, mean = 1:10, sd = 1)
generate(my_dist, 10) # lapply(1:10, function(mu) rnorm(10, mean = mu, sd = 1))
```

You can also compute intervals using `hilo()`

```{r hilo}
hilo(my_dist, 0.95)
```

Additionally, some distributions may support other methods such as mathematical operations and summary measures.

```{r math}
my_dist
my_dist*3 + 2
mean(my_dist)
variance(my_dist)
```
